---
title: "BST225"
output: pdf_document
---

# HW5
## 1a)
Calculate sample size using all these values, change values for one factor at one time and keep other factors the sampe; and make a sample size analysis table; then summarize how difference between u1 and u0, sigma and power affect sample size
```{r}
for (power in c(0.95,0.9,0.85,0.8))
{
  for (sigma_1 in c(2.2,2.0,1.8))
  {
    for (sigma_2 in c(1.8))
    {
      for (mu1_min_mu2 in c(1.5,1.0,0.5))
      {
        
        n <- ((sigma_1^2 + sigma_2^2)/(mu1_min_mu2)^2)*(qnorm(0.05/2)+qnorm(1-power))^2
        print (sprintf("Power: %s  Sigma1: %s  Sigma2: %s  Mu1-Mu2: %s  Sample size: %s",  power, sigma_1, sigma_2, mu1_min_mu2, round(n)))
      }
    }
  }
}
  
```
- As difference between mu1 and mu2 decrease the sample size needs to increase. 
- As the pooled variance decreases then the necessary sample size will also decrease. 
- As the power decreases the sample size decreases. 


## 1b) Is this what is wanted?Z
300 and 700 as n1 and n2
Calculate power using parameters values and make a power analysis table and summarize the effect of the parameters. 

### For n2, For n1 (same result as n2)

```{r}
n1 <- 300
n2 <- 700
k <- n1/n2

for (sigma_2 in c(2.2,2.0,1.8))
{
  for (mu1_min_mu2 in c(1.5,1.0,0.5))
  {  
    q_norm_power <- sqrt(n2/((sigma_2^2*(1+1/k))/(mu1_min_mu2)^2)) + qnorm(0.05/2)
    print (sprintf("Power: %s  Sigma1: %s  Sigma2: %s  Mu1-Mu2: %s ",  pnorm(q_norm_power), sigma_1, sigma_2, mu1_min_mu2))
  }
}  



```
- As the sample size increase so does the power
- As the difference between the two groups incrases then the power increases
- As the pooled variance between the two groups increases the the power decreases




## 2)
Test non-inferiority with two different margins and study how the values of margins affect the sample size. 
```{r}
for (power in c(0.9,0.8))
{
  for (sigma_1 in c(2.2))
  {
    for (sigma_2 in c(1.8))
    {
      for (mu1_min_mu2 in c(1.0,0.5))
      {
        for (margin in c(1.1,0.4))
        {
          n <- ((sigma_1^2 + sigma_2^2)/(mu1_min_mu2-margin)^2)*(qnorm(0.05/2)+qnorm(1-power))^2
          print (sprintf("Power: %s  Sigma1: %s  Sigma2: %s  Mu1-Mu2: %s  Sample size: %s  Margin: %s",  power, sigma_1, sigma_2, mu1_min_mu2, round(n), margin))
        }
      }
    }
  }
}
```
- As the difference between mu1-mu2 and margin decreases then the necessary sample size to show inferiority increases.

Test equivalence with two different margins and study how the values of margins affect the sample size. 
```{r}
for (power in c(0.9,0.8))
{
  for (sigma_1 in c(2.2))
  {
    for (sigma_2 in c(1.8))
    {
      for (mu1_min_mu2 in c(1.0,0.5))
      {
        for (margin in c(1.1,0.4))
        {
          n <- ((sigma_1^2 + sigma_2^2)/(mu1_min_mu2-margin)^2)*(qnorm(0.05/2)+qnorm((1-power)/2))^2
          print (sprintf("Power: %s  Sigma1: %s  Sigma2: %s  Mu1-Mu2: %s  Sample size: %s  Margin: %s",  power, sigma_1, sigma_2, mu1_min_mu2, round(n), margin))
        }
      }
    }
  }
}
```
- As the difference between abs(mu1-mu2) and margin decreases then the necessary sample size to show equivalence increases.




# HW6

## 1a)
Calculate sample size using all these values, change values for one factor at one time and keep other factors the sampe; and make a sample size analysis table; then summarize how difference between p1 and p2, sigma and power affect sample size
```{r}
for (power in c(0.95,0.9,0.85,0.8))
{
  for (p_1 in c(0.85))
  {
    for (p_2 in c(0.8,0.7,0.6))
    {
        n <- ((p_1*(1-p_1) + p_2*(1-p_2))/(p_1-p_2)^2)*(qnorm(0.05/2)+qnorm(1-power))^2
        print (sprintf("Power: %s  P1: %s  P2: %s  Sample size: %s",  power, p_1, p_2, round(n)))
    
    }
  }
}
  
```
- As power decreases so does the necessary sample size
- As the difference between p_1 and p_2 increases the sample size decreases


## 1b)
Sample size = 90,170  for each arm, calculate power using the above parameter values, change values for one factor at one time and keep other factors the same. Make a power analysis table and summarize how the difference between p1 and p2, and sample size affect power. 

```{r}
for (n_1 in c(80,90))
{
  for (n_2 in c(170,180,190))
  {
    for (p_1 in c(0.85))
    {
      for (p_2 in c(0.8,0.7,0.6))
      {
          k <- n_1/n_2
          q_norm_power <- sqrt(n/((((p_1*(1-p_1))/k + p_2*(1-p_2))/(p_1-p_2)^2))) + qnorm(0.05/2)
          print (sprintf("Power: %s  P1: %s  P2: %s  N1: %s  N2: %s",  pnorm(q_norm_power), p_1, p_2, n_1, n_2))
      
      }
    }    
  }
}
  
```

- As the difffernce between p_1 and p_2 increases then so does the power
- As the sample size increases for n_1 then the power increases.
- As the sample size incrase for n_2 then the power increases






# HW7 
Trial to compare a new therapy with a routine bath care in terms of the time to infection.
- Hazard ratio (HR) of 1.5, 2.5, (beta=log(HR)), respectively, for routine bathing care/test therapy is considered of clinical importance.
- 60%, 80% of patients’ infection may be observed, respectively (=d)
- n=n1=n2 (p1=p2=0.5), equal size treatment groups
- Significance level: a=0.05

## 1A)
If Power = 0.8, b=1-0.8=0.2, calculate sample size for above various HRs and d’s.
```{r}
for (d in c(0.6, 0.8))
{
  for (hr in c(1.5, 2.0, 2.5))
  {
    n <- (qnorm(0.05/2) + qnorm(1-0.8))^2/(log(hr)^2*0.5*0.5*d)
    print(sprintf("HR: %s  Patiends observed(d): %s  Sample size: %s", hr, d, n))
  }    
}

```


## 1B)
If n1 = n2 = 20, 50, respectively, calculate power for above various HRs and d’s.
```{r}
for (n in c(20,50))
{
  for (d in c(0.6, 0.8))
  {
    for (hr in c(1.5, 2.5))
    {
      q_norm_power <- - sqrt((n*(log(hr)^2*0.5*0.5*d))) - (qnorm(0.05/2))
      print(sprintf("HR: %s  Patiends observed(d): %s  Sample size: %s  Power: %s", hr, d, n, pnorm(q_norm_power)))
    }    
  }
}

```











