---
title: "HW#4"
author: "Keith Mitchell"
date: "2/20/2020"
output:
  pdf_document: default
  word_document: default
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Partial code credit to Prof Li (UC Davis), Weiping Zhang(USTC)


## Number 2 
### Part 1
#### two-sample Hotelling's T2 test  -------

```{r}
# now we perform the two-sample Hotelling T^2-test
n<-c(8,8)
p<-3
xmean1<-c(82,81,80)
xmean2<-c(80,81,82)
d<-xmean1-xmean2
solve(matrix(c(0.5,-0.25,-0.25,0.5), nrow=2))
S1<-matrix(c(9,5,5,5,9,5,5,5,9), nrow=3)
S2<-matrix(c(9,5,5,5,9,5,5,5,9), nrow=3)
Sp<-((n[1]-1)*S1+(n[2]-1)*S2)/(sum(n)-2)
Sp
t2 <- t(d)%*%solve(sum(1/n)*Sp)%*%d
alpha<-0.05
t2
cval <- (sum(n)-2)*p/(sum(n)-p-1)*qf(1-alpha,p,sum(n)-p-1)
cval
```
#### Since T^2 = 12 > 6.76 the null hypothesis is rejected at 5% level of significance.


### Part 2
#### Confidence Region

```{r}
es<-eigen(sum(1/n)*Sp)
e1<-es$vec %*% diag(sqrt(es$val))
r1<-sqrt(cval)
theta<-seq(0,2*pi,len=250)
v1<-cbind(r1*cos(theta), r1*sin(theta))
pts<-t(d-(e1%*%t(v1)))
plot(pts,type="l",main="Confidence Region for Bivariate Normal",xlab=expression(paste(mu, "(1)")), ylab=expression(paste(mu, "(2)")),asp=1)
segments(0,d[2],d[1],d[2],lty=2) # highlight the center
segments(d[1],0,d[1],d[2],lty=2)

th2<-c(0,pi/2,pi,3*pi/2,2*pi)   #adding the axis
v2<-cbind(r1*cos(th2), r1*sin(th2))
pts2<-t(d-(e1%*%t(v2)))
segments(pts2[3,1],pts2[3,2],pts2[1,1],pts2[1,2],lty=3)  
segments(pts2[2,1],pts2[2,2],pts2[4,1],pts2[4,2],lty=3)

# since we reject the null, we use the simultaneous confidence intervals
# to check the significant components
```

### Part 3
#### Simultaneous confidence intervals
```{r}
wd<-sqrt(cval*diag(Sp)*sum(1/n))
Cis<-cbind(d-wd,d+wd)
# 95% simultaneous confidence interval
Cis
#plot(Cis[1][1]:0, 1:10, type="l", lty=2)
```
### Part 4
#### Bonferroni simultaneous confidence intervals
```{r}
wd.b<- 2.977*sqrt(diag(Sp)*sum(1/n))
sqrt(diag(Sp)*sum(1/n))
qt(1-alpha/(2*p),n[1]+n[2]-2)
Cis.b<-cbind(d-wd.b,d+wd.b)
# 95% Bonferroni simultaneous confidence interval
Cis.b
# both component-wise simultaneous confidence intervals do not contain 0, so they have significant differences.
```


## Number 3
### Part 1
#### two-sample Hotelling's T2 test  -------
```{r}
c <- matrix(c(-1,0,1,-1,0,1), nrow=2)
tran_c <- t(c)
s_1 <- matrix(c(16,8,8,8,16,8,8,8,16), nrow=3)
corrected_s <- c%*%s_1%*%tran_c
```


```{r}
# now we perform the two-sample Hotelling T^2-test
n<-c(18,18)
p<-2
xmean1<-c(0,0)
xmean2<-c(2,2)
d<-xmean1-xmean2
S1<-corrected_s
S2<- corrected_s
Sp<-((n[1]-1)*S1+(n[2]-1)*S2)/(sum(n)-2)
t2 <- t(d)%*%solve(sum(1/n)*Sp)%*%d
t2

alpha<-0.05
cval <- (sum(n)-2)*p/(sum(n)-p-1)*qf(1-alpha,p,sum(n)-p-1)
cval
```
#### Since T^2 = 9 > 6.76 the null hypothesis is rejected at 5% level of significance. 

### Part 2
#### Confidence Region
```{r}
es<-eigen(sum(1/n)*Sp)
e1<-es$vec %*% diag(sqrt(es$val))
r1<-sqrt(cval)
theta<-seq(0,2*pi,len=250)
v1<-cbind(r1*cos(theta), r1*sin(theta))
pts<-t(d-(e1%*%t(v1)))
plot(pts,type="l",main="Confidence Region for Bivariate Normal",xlab=expression(paste(mu, "(1)")), ylab=expression(paste(mu, "(2)")),asp=1)
segments(0,d[2],d[1],d[2],lty=2) # highlight the center
segments(d[1],0,d[1],d[2],lty=2)

th2<-c(0,pi/2,pi,3*pi/2,2*pi)   #adding the axis
v2<-cbind(r1*cos(th2), r1*sin(th2))
pts2<-t(d-(e1%*%t(v2)))
segments(pts2[3,1],pts2[3,2],pts2[1,1],pts2[1,2],lty=3)  
segments(pts2[2,1],pts2[2,2],pts2[4,1],pts2[4,2],lty=3)

# since we reject the null, we use the simultaneous confidence intervals
# to check the significant components
```
### Part 3
#### Simultaneous confidence intervals

```{r}

wd<-sqrt(cval*diag(Sp)*sum(1/n))
Cis<-cbind(d-wd,d+wd)

# 95% simultaneous confidence interval
Cis
#plot(Cis[1][1]:0, 1:10, type="l", lty=2)
```
### Part 4
#### Bonferroni simultaneous confidence intervals
```{r}
wd.b<- qt(1-alpha/(2*p),n[1]+n[2]-2) *sqrt(diag(Sp)*sum(1/n))
Cis.b<-cbind(d-wd.b,d+wd.b)
# 95% Bonferroni simultaneous confidence interval
Cis.b
# both component-wise simultaneous confidence intervals do not contain 0, so they have significant differences.
```


## Number 4
#### two-sample Hotelling's T2 test
```{r}
# now we perform the two-sample Hotelling T^2-test
n<-c(45,55)
p<-2
xmean1<-c(204.4,556.6)
xmean2<-c(130,355)
d<-xmean1-xmean2
S1 <- matrix(c(13825.3, 23823.4, 23823.4, 73107.4), nrow=2)
S2 <- matrix(c(8632.0, 19616.7, 19616.7, 55964.5), nrow=2)

Sp<-((n[1]-1)*S1+(n[2]-1)*S2)/(sum(n)-2)
t2 <- t(d)%*%solve(sum(1/n)*Sp)%*%d
t2

alpha<-0.05
cval <- (sum(n)-2)*p/(sum(n)-p-1)*qf(1-alpha,p,sum(n)-p-1)
cval
```

#### Since T^2 = 16.06 > 6.24 the null hypothesis is rejected at 5% level of significance. 

```{r}
a <- solve(Sp)%*%d
a
```

#### Where a is the linear combination of mean components most responsbile for the rejection of the null hypothesis.  