---
title: "HW6"
author: "Keith Mitchell"
date: "3/7/2020"
output: pdf_document
---

# Question 1:
## a) Give the formula for the first and second principle components 
```{r}
X <- as.matrix(cbind(c(2,3,4,0,-1,-2,-3), c(1,1,1,0,-1,-1,-1)))
Z <- as.matrix(cbind(c(7.17, 2.5), c(2.5, 1)))
Z
X
```

```{r}
help <- (1/6)*t(X)%*%X
help
cov <- cov(X)
cov
cor <- cor(X)
cor
ev_z <- eigen(cor)
ev_z$values
ev_z$vectors

ev <- eigen(cov)
ev$values
ev$vectors
```


### So the first principle component is the eigen vector corresponding to the largest eigenvalue which is the first column above
```{r}
ev$vectors[,1]
```
### So the second principle component is the eigen vector corresponding to the second largest eigenvalue which is the second column above
```{r}
ev$vectors[,2]
```

## b) Determine the proportion of total sample variance due to the first sample principal component.
```{r}
ev$values[1]/sum(ev$values)
```

## c) Compare the contributions of the two variates to the determination of the first sample principal component based on loadings
- here we see that the first value of the eigen vector is the contribution of the first variable towards the principle component while the second value is the contribution of the second variable to the principle component
```{r}
ev$vectors[,1]
```


## d) Compare the contributions of the two variates to the determination of the first sample principal component based on sample correlations

```{r}
corr_1 <- ev$vectors[,1][1]*sqrt(ev$values[1]/cov[,1][1])
corr_1
corr_2 <- ev$vectors[,1][2]*sqrt(ev$values[1]/cov[,2][2])
corr_2
```


```{r}
cor <- cor(X)
cor
ev <- eigen(cor)
ev$values
ev$vectors
```



## e) Redo (a)-(d) on the standardized dataset. 
### a) Give the formula for the first and second principle components 
```{r}
X_stan <- scale(X)
X_stan[is.nan(X_stan)] <- 0
X_stan
```


```{r}
cov <- cov(X_stan)
cov
ev <- eigen(cov)
ev$values
ev$vectors
```

#### So the first principle component is the eigen vector corresponding to the largest eigenvalue which is the first column above
```{r}
ev$vectors[,1]
```
#### So the second principle component is the eigen vector corresponding to the second largest eigenvalue which is the second column above
```{r}
ev$vectors[,2]
```

### b) Determine the proportion of total sample variance due to the first sample principal component.
```{r}
ev$values[1]/sum(ev$values)
```

### c) Compare the contributions of the two variates to the determination of the first sample principal component based on loadings
- here we see that the first value of the eigen vector is the contribution of the first variable towards the principle component while the second value is the contribution of the second variable to the principle component
```{r}
ev$vectors[,1]
```


### d) Compare the contributions of the two variates to the determination of the first sample principal component based on sample correlations

```{r}
ev$vectors[,1][1]
ev$vectors[,1][2]
corr_1 <- ev$vectors[,1][1]*sqrt(ev$values[1])
corr_1
corr_2 <- ev$vectors[,1][2]*sqrt(ev$values[1])
corr_2
```








# Question 3:
## Consider the air polution in table 1.5. Summarize the data in fewer the p=7 dimensions if possible. Conduct a PCA of the data using both the covariance matrix S and the correlation matrix R. What have you learned? Does it make any difference which matrix is chose for analysis? Can the data be summarized in thre or fewer dimensions? Can you interpret the principal components?

```{r}
data <- read.table("T1-5.DAT", 
           header=FALSE)
data <- as.matrix(data)

```

```{r}
cov <- cov(data)
cor <- cor(data)

cov
cor
```


## The eigenvalues and vectors based on the covariance matrix are:

```{r}
ev <- eigen(cov)
ev$values
ev$vectors
```

## Then to calculate the proportion of variance in the first two pcs
```{r}
(sum(ev$values[1:2]))/sum(ev$values)
```

## So the first and second principal components summarize 95.4% of the variation in the data based on the covariance matrix. 


## In comparison, the eigenvalues and vectors based on the correlation matrix are:
```{r}
ev <- eigen(cor)
sum(ev$values)
ev$values
ev$vectors
```


## Then to calculate the proportion of variance in the first two pcs and 3 pcs
```{r}
(sum(ev$values[1:2]))/sum(ev$values)
(sum(ev$values[1:3]))/sum(ev$values)

```

## So the first and second principal components summarize 53.2% of the variation in the data based on the corrlation matrix. The first 3 pcs are ~70%.


## Based on these results the choice of the covariance vs the correlation matrix makes a difference. The correlation matrix can be summarized in the first 3 principal components where as the covariance matrix can be summarized in the first two principla components fairly effectively. The data can be summarized in 3 or fewer dimensions, but it is not very effective in comparison.







# Question 7
```{r}
data <- read.table("T1-6.DAT", 
           header=FALSE, colClasses = c("integer", "double", "double", "double", "double", "factor"))
colnames(data) <- c("Age", "S1L+S1R", "S1L-S1R", "S2L+S2R", "S2L-S2R", "Group")
levels(data$Group)[levels(data$Group)=="0"] <- "Non-MS"
levels(data$Group)[levels(data$Group)=="1"] <- "MS"
summary(data)
```
```{r}
group_1 <- data[data$Group == "Non-MS",]

print("Summary and dimensions of the first group of individuals.")
summary(group_1)
dim(group_1)
group_2 <- data[data$Group == "MS",]

print("Summary and dimensions of the second group of individuals.")
summary(group_2)
dim(group_2)
X <- as.matrix(cbind(data[1:5]))
group_list <- data[6]
X_1 <- as.matrix(cbind(group_1[1:5]))
X_2 <- as.matrix(cbind(group_2[1:5]))
```

## Some general setup for fishers rule
```{r}
S1 <- cov(X_1)
S2 <- cov(X_2)
n1 <- nrow(X_1)
n2 <- nrow(X_2)
n<-c(n1,n2)
xmean1 <- colMeans(X_1)
xmean2 <- colMeans(X_2)
d<-xmean1-xmean2
Sp<-((n[1]-1)*S1+(n[2]-1)*S2)/(sum(n)-2)
Sp
```

##  Construct Fisher’s rule. Moreover,calculate theapparent error rate.
```{r}
w <- solve(Sp)%*%(xmean1-xmean2)
correct <- 0
incorrect <- 0
for (val in 1:nrow(X))
{
  row <- X[val,]
  group <- group_list[val,]
  #print("-----")
  #print(group)
  left_side <- t(w)%*%row
  right_side <- 0.5*t(w)%*%(xmean1+xmean2)
  #print(left_side)
  #print(right_side)
  if (left_side >= right_side){
    if (group == "Non-MS"){
      #print("Correct, should be Non-mS")
      correct <- correct + 1
    }
    else{
      #print("incorrect, should be MS")
      incorrect <- incorrect + 1
    }
  }
  else{
    if (group == "Non-MS"){
      #print("incorrect, should be MS")
      incorrect <- incorrect + 1
    }
    else{
      #print("Correct, should be ms")
      correct <- correct + 1
    }
  }
}
print("Error rate for Lachenbruch’s holdout:")
incorrect/(nrow(X))
print("Number correct:")
correct
print("Number incorrect:")
incorrect
```

## Finally the expected actual error rate by Lachenbruch’s holdout.

```{r}
# this was fun :)
incorrect <- 0
correct <- 0
for (val in 1:nrow(data))
{
  # first we want to take out a row in the dataframe
  data_check <- data[c(val),]
  data_seg <- data[-c(val),]
  
  # then we want to seperate the data into the two groups
  group_1 <- data_seg[data_seg$Group == "Non-MS",]
  group_2 <- data_seg[data_seg$Group == "MS",]
  
  # we also want a matrix of the data and group_list calculate w based on s_pooled (since size of xmean1/xmean2 and n1/n2 are going to change)
  X_1 <- as.matrix(cbind(group_1[1:5]))
  X_2 <- as.matrix(cbind(group_2[1:5]))
  S1 <- cov(X_1)
  S2 <- cov(X_2)
  n1 <- nrow(X_1)
  n2 <- nrow(X_2)
  n<-c(n1,n2)

  xmean1 <- colMeans(X_1)
  xmean2 <- colMeans(X_2)
  d<-xmean1-xmean2
  Sp<-((n[1]-1)*S1+(n[2]-1)*S2)/(sum(n)-2)
  w <- solve(Sp)%*%(xmean1-xmean2)
  
  # lets set row to the row we took out of the dataframe (to match our code from earlier) probably should make this a function call but whatever
  row <- as.matrix(cbind(data_check[1:5]))[1,]
  group <- data_check[6][1,]
  left_side <- t(w)%*%row
  right_side <- 0.5*t(w)%*%(xmean1+xmean2)
  if (left_side >= right_side){
    if (group == "Non-MS"){
      #print("Correct, should be Non-mS")
      correct <- correct + 1
    }
    else{
      #print("incorrect, should be MS")
      incorrect <- incorrect + 1
    }
  }
  else{
    if (group == "Non-MS"){
      #print("incorrect, should be MS")
      incorrect <- incorrect + 1
    }
    else{
      #print("Correct, should be ms")
      correct <- correct + 1
    }
  }
}
apparent_error_rate <- incorrect/(incorrect+correct)
print("Error rate for Lachenbruch’s holdout:")
incorrect/(nrow(X))
print("Number correct:")
correct
print("Number incorrect:")
incorrect
```











